Hexagonal binning algorithm

### 2024-05-21 supplementary/hexagon_binning_algorithm.qmd
Updated figures in hexagon algorithm doc

```{r}
hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2 - 0.05), fill = "white", color = "black", alpha = 0, linewidth = 1) +
    geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
    geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
    annotate("text", x=-0.18, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
  coord_equal()
```

```{r}
#| label: fig-hex-param
#| fig-cap: "Notations for hexagonal grid configurations."
#| out-height: 30%
#| fig-pos: H

@@

assign_data_vis2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd")
```

```{r}
#| fig-width: 9
#| fig-height: 6
assign_data_vis1 + assign_data_vis2 +
  plot_layout(ncol = 2)
```

```{r}

## To generate embedding data
emb_df <- tibble::tibble(emb1 = 0.37, emb2 = 0.3) |>
  add_row(emb1 = 0.526, emb2 = 1.1) |>
  add_row(emb1 = 0.735, emb2 = 1.1) |>
  add_row(emb1 = 0.631, emb2 = 1.22) |>
  add_row(emb1 = 0.735, emb2 = 1.16) |>
  add_row(emb1 = 0.683, emb2 = 1.19) |>
  add_row(emb1 = 0.735, emb2 = 1.04) |>
  add_row(emb1 = 0.631, emb2 = 0.981) |>
  add_row(emb1 = 0.683, emb2 = 1.0105) |>
  add_row(emb1 = 0.526, emb2 = 1.04) |>
  add_row(emb1 = 0.5785, emb2 = 1.0105) |> 
  add_row(emb1 = 0.526, emb2 = 1.16) |>
  add_row(emb1 = 0.5785, emb2 = 1.19)

## Data to generate circle
circles <- data.frame(
  x = c(0.318, 0.422),
  y = c(0.197,0.378),
  r = rep(0.1205264,2)
)

assign_data_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 2) +
  geom_point(data = emb_df, aes(x = emb1, y = emb2), color = "#253494", size = 0.5) +
  geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
  geom_segment(x = 0.37, y = 0.3, xend = 0.422, yend = 0.378, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.526, y = 1.1, xend = 0.45, yend = 1.1, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.735, y = 1.1, xend = 0.66, yend = 1.1, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.631, y = 1.22, xend = 0.631, yend = 1.15, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.735, y = 1.16, xend = 0.65, yend = 1.12, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.735, y = 1.04, xend = 0.735, yend = 0.95, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.683, y = 1.19, xend = 0.65, yend = 1.12, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.631, y = 0.981, xend = 0.55, yend = 0.95, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.683, y = 1.0105, xend = 0.7, yend = 0.95, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.526, y = 1.04, xend = 0.526, yend = 0.95, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.5785, y = 1.0105, xend = 0.536, yend = 0.97, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.526, y = 1.16, xend = 0.49, yend = 1.12, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  geom_segment(x = 0.5785, y = 1.19, xend = 0.58, yend = 1.12, 
               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
  coord_equal()

@@

```{r}
#| label: fig-assign-data
#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."

assign_data_vis
```

>>>>

```{r}
hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354", size = 0.9) +
    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2 - 0.05), fill = "white", color = "black", alpha = 0, linewidth = 0.7) +
    geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
    geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
    annotate("text", x=-0.18, y=-0.45, label= "(s1, s2)", color = "black") ## starting point
```

```{r}
#| label: fig-hex-param
#| fig-cap: "Notations for hexagonal grid configurations."
#| fig-pos: H
#| fig-width: 2
#| fig-height: 3

@@

emb_df2 <- tibble(emb1 = 0.161, emb2 = 0.7395) |>
  add_row(emb1 = 0.683, emb2 = 0.7395) |>
  add_row(emb1 = 0.2915, emb2 = 0.965) |>
  add_row(emb1 = 0.5525, emb2 = 0.965) |>
  add_row(emb1 = 0.2915, emb2 = 0.5135) |>
  add_row(emb1 = 0.5525, emb2 = 0.5135) |>
  add_row(emb1 = 0.422, emb2 = 1.04) |>
  add_row(emb1 = 0.161, emb2 = 0.890) |>
  add_row(emb1 = 0.161, emb2 = 0.589) |>
  add_row(emb1 = 0.422, emb2 = 0.438) |>
  add_row(emb1 = 0.683, emb2 = 0.589) |>
  add_row(emb1 = 0.683, emb2 = 0.890)

assign_data_vis2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
  geom_point(data = emb_df2, aes(x = emb1, y = emb2), color = "black", size = 0.7) +
  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
  geom_segment(x = 0.161, y = 0.7395, xend = -0.06, yend = 0.740, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.683, y = 0.7395, xend = 0.45, yend = 0.740, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.2915, y = 0.965, xend = 0.43, yend = 0.79, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.5525, y = 0.965, xend = 0.44, yend = 0.8, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.2915, y = 0.5135, xend = 0.18, yend = 0.32, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.5525, y = 0.5135, xend = 0.67, yend = 0.32, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.422, y = 1.04, xend = 0.42, yend = 0.85, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.161, y = 0.890, xend = -0.02, yend = 0.8, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
  geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")

@@

```{r}
#| label: fig-assign-data
#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."
#| fig-width: 9
#| fig-height: 6
assign_data_vis1 + assign_data_vis2 +
  plot_layout(ncol = 2)
```
